---
title: Basic concepts of quasispecies simulation
output: 
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
      sharing:
        facebook: no
        twitter: no
      fontsettings:
        theme: sepia
---

```{r, include=FALSE}
library(yasss)
library(knitr)
library(rmarkdown)
library(ggthemr)

knitr::opts_chunk$set(echo = FALSE)
ggthemr('lilac')

c_theme <- theme_get()
c_theme <- c_theme + theme(plot.background = element_rect(fill = "#F3EACB"))
old_theme <- theme_set(c_theme)
```

```{r , include = FALSE}
source('../utilities/reporting_utilities.R')
```

# Simple first simulation

A first simulation to demonstrate the process and the measuring tools. Table \@ref(tab:sim-pop-config) shows the configuration that was used to call `sim_pop` from the `yasss` package. A high level summary of the entire genealogy is provided in Table \@ref(tab:first-sum-tab).

```{r sim-pop-config, results = 'asis'}
args <- list(
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 9,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 0.1)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL)
)

x <- do.call(sim_pop, args)

kable(
  sim_pop_arg_formatter(args),
  booktabs = TRUE,
  caption = "Configuration with which sim_pop was called."
)
```

```{r first-sum-tab, results = 'asis'}
g_sum_tab <- genealogy_summary_table(x)

kable(
  g_sum_tab,
  booktabs = TRUE,
  caption = "Individuals per generation and the average distance to the original ancestor for the generation."
)
```

Our primary interested is restricted to the final generation which corresponds to the generation that we will obtain from sequencing a sample.
Hence we need to capture some of the key information about the last generation and figure out ways to compare it to other simulations.
Three key ways to look at this data is:

- The average of all the pairwise distances between the sequences.
- A density plot of the distribution of all the pairwise distances.
- The deciles / quartiles / some other set of percentiles of the distribution.

For now we will present all three. Once we have some solid results to argue about we can look at dropping some of these metrics.

```{r, include = FALSE}
last_gen <-
x %>%
  filter(gen_num == max(gen_num))

dmat <- stringdistmatrix(last_gen$the_seq, method = 'hamming')
avg_hd <- mean(dmat)
dvec <- as.numeric(dmat)/min(nchar(last_gen$the_seq))
dd.f <- data.frame(dists = dvec)
```

The average pairwise HD is **`r round(avg_hd, 2)`** and normalized by the length of the sequence it is **`r round(avg_hd/min(nchar(last_gen$the_seq)), 3)`**. The distribution is shown in Figure \@ref(fig:first-density-plot) and the deciles are listed in Table \@ref(tab:first-deciles).

```{r first-density-plot, fig.cap='Density plot of the pariwise distances in the last generation', results = 'asis'}
n_uniq <- nrow(unique(dd.f))
range_length <- max(dd.f$dists) - min(dd.f$dists)
bw <- range_length / min((n_uniq*1.75), 200)
ggplot(dd.f, aes(x = dists)) + geom_density(bw = bw)
```

```{r first-deciles, results = 'asis'}
deciles <-
data.frame(Decile = paste((0:10)*10, "%", sep = ''),
           Value = quantile(dvec, (0:10)/10))
row.names(deciles) <- NULL
kable(
  deciles,
  booktabs = TRUE,
  caption = "The deciles of the pairwise HDs"
)
```

# More realistic mu

Note that the extremely high value of mu (0.1) produces sequences that rapidly diverge radically from the ancestor. To make the simulation a little more realistic, drop the mutation rate down to 1 in 1000 and try again.

```{r sim-mu-1-1000, results = 'asis'}
args <- list(
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 9,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 1/1e3)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL)
)

x <- do.call(sim_pop, args)

sim_result <- list(args = args,
                   genealogy = x)
```

```{r test-child, child = '../utilities/single_sim_chunk.Rmd'}
```
