---
title: Basic concepts of quasispecies simulation
output: 
  html_document:
    toc: false
    theme: cerulean
---

```{r, include=FALSE}
library(yasss)
library(knitr)
library(rmarkdown)
library(tidyr)

knitr::opts_chunk$set(echo = FALSE)
options(scipen = 99)

report_start_time <- proc.time()
```

```{r , include = FALSE}
source('../utilities/reporting_utilities.R')
```

```{r, eval = FALSE}
yasss:::restart_r() 
```


# Simple first simulation

In early versions of `yasss`, `sim_pop` was the main simulation function. It
has been wrapped by the `sim_proc_many_pops` function in later versions. This
report was originally written for use with the `sim_pop` function. I have just
updated the code blocks to use `sim_proc_many_pops`. So instead of manually
stepping through the processing steps in this document, it now just retrieves
the values from the return value of `sim_proc_many_pops` and presents the
findings. Thus, this report can be seen as partial documentation of the
internals of `sim_proc_many_pops`. The writeup in this report was only
annotated to indicate where it is not really accurate anymore. Such
annotations are indicated by **NEW:**.

A first simulation to demonstrate the process and the measuring tools. The
following table shows the configuration that was used to call `sim_pop` from
the `yasss` package. A high level summary of the entire genealogy is provided
in the second table. **NEW:** Such a configuration for a single simulation is
stored in an `arg_set`, and the results of `sim_pop` are extracted from a
`sim_proc_many_pops` return value by ensuring that `sim_proc_many_pops` is
called with the option `output_genealogy = 'full'` and extracting the
sub-elements of the `all_genealogies` element of `sim_proc_many_pops` return
value.

```{r sim-pop-config, results = 'asis'}
arg_set1 <- list(
  label = 'A',
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 9,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 0.1)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL),
  required_fitness = 0
)

many_pops <- sim_proc_many_pops(list(arg_set1), output_genealogy = 'full', output_dmat = TRUE)
x <- many_pops$all_genealogies[[1]]

kable(
  sim_pop_arg_formatter(arg_set1),
  booktabs = TRUE,
  caption = "Configuration with which sim_pop was called."
)
```

```{r first-sum-tab, results = 'asis'}
g_sum_tab <- genealogy_summary_table(x)

kable(
  g_sum_tab,
  booktabs = TRUE,
  caption = "Individuals per generation and the average distance to the original ancestor for the generation."
)
```

Our primary interested is restricted to the final generation which corresponds
to the generation that we will obtain from sequencing a sample.  Hence we need
to capture some of the key information about the last generation and figure
out ways to compare it to other simulations.  Three key ways to look at this
data is:

- The average of all the pairwise distances between the sequences.
- A density plot of the distribution of all the pairwise distances.
- The percentiles of the distribution.

**NEW:** The distance matrices can be output by `sim_proc_many_pops` by
setting the `output_dmat = TRUE` option. The summaries of the distance
matrices can be extracted from the individual `dsum`s stored in the
`dcollection` element of the return value from `sim_proc_many_pops`.

```{r, include = FALSE}
last_gen <-
x %>%
  filter(gen_num == max(gen_num))

dmat <- many_pops$all_dmats[[1]]$dmat
avg_hd <- mean(dmat)
stopifnot(avg_hd == many_pops$dcollection[[1]]$avg_hd)
dvec <- as.numeric(dmat)/min(nchar(last_gen$the_seq))
dd.f <- data.frame(dists = dvec)
```

The average pairwise HD is **`r round(avg_hd, 2)`** and normalized by the
length of the sequence it is **`r round(avg_hd/min(nchar(last_gen$the_seq)),
3)`**. The distribution is shown in the figure below and the deciles are
listed in a table below the figure. Note that the bandwidth has to be tweaked
because mutations are rare events thus the distribution is fairly discrete.

```{r first-density-plot, fig.cap='Density plot of the pariwise distances in the last generation', results = 'asis'}
n_uniq <- nrow(unique(dd.f))
range_length <- max(dd.f$dists) - min(dd.f$dists)
bw <- range_length / min((n_uniq*1.75), 200)
ggplot(dd.f, aes(x = dists)) + geom_density(bw = bw)
```

```{r first-deciles, results = 'asis'}
deciles <-
data.frame(Decile = paste((0:10)*10, "%", sep = ''),
           Value = quantile(dvec, (0:10)/10))
row.names(deciles) <- NULL
kable(
  deciles,
  booktabs = TRUE,
  caption = "The deciles of the pairwise HDs"
)
```

# More realistic mu

Note that the extremely high value of mu (0.1) produces sequences that rapidly diverge radically from the ancestor. To make the simulation a little more realistic, drop the mutation rate down to 1 in 1000 and try again.

```{r sim-mu-1-1000, results = 'asis'}
arg_set2 <- list(
  label = 'A',
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 9,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 1/1e3)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL),
  required_fitness = 0
)

many_pops <- sim_proc_many_pops(list(arg_set2), output_genealogy = 'full', output_dmat = TRUE)
sim_result <- list()
sim_result$genealogy <- many_pops$all_genealogies[[1]]
```

```{r test-child, child = '../utilities/single_sim_chunk.Rmd'}
```

# Small number of simulations

The goal is to compare different simulations to each other. First, we will start with a small number of simulations that we can still pay attention to each individual simulation.

```{r, include = FALSE}
arg_set3 <- list(
  label = 'A',
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 9,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 1/500)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL),
  required_fitness = 0
)

many_pops <- sim_proc_many_pops(list(arg_set3), output_genealogy = 'full', output_dmat = TRUE,
                                n_sims = 10)
x <- rbindlist(many_pops$all_genealogies)
all_sim_results <- data.frame(x)


#all_sim_results <- NULL
#
#i <- 1
#for (i in 1:10){
#  x <- do.call(sim_pop, args)
#  original_cols <- names(x)
#  x$sim_id <- i
#  x <- x %>% select(sim_id, original_cols)
#  all_sim_results <- rbind(all_sim_results, x)
#}
```

```{r test-child, child = '../utilities/few_sim_chunk.Rmd'}
```


