---
title: Basic concepts of quasispecies simulation
output: 
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
      sharing:
        facebook: no
        twitter: no
      fontsettings:
        theme: sepia
---

```{r, include=FALSE}
library(yasss)
library(knitr)
library(rmarkdown)
library(ggthemr)

knitr::opts_chunk$set(echo = FALSE)
ggthemr('lilac')

c_theme <- theme_get()
c_theme <- c_theme + theme(plot.background = element_rect(fill = "#F3EACB"))
old_theme <- theme_set(c_theme)
```

```{r , include = FALSE}
source('../utilities/reporting_utilities.R')
```

# Simple first simulation

A first simulation to demonstrate the process and the measuring tools. Table \@ref(tab:sim-pop-config) shows the configuration that was used to call `sim_pop` from the `yasss` package. A high level summary of the entire genealogy is provided in Table \@ref(tab:first-sum-tab).

```{r sim-pop-config, results = 'asis'}
args <- list(
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 9,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 0.1)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL)
)

x <- do.call(sim_pop, args)

kable(
  sim_pop_arg_formatter(args),
  booktabs = TRUE,
  caption = "Configuration with which sim_pop was called."
)
```

```{r first-sum-tab, results = 'asis'}
g_sum_tab <- genealogy_summary_table(x)

kable(
  g_sum_tab,
  booktabs = TRUE,
  caption = "Individuals per generation and the average distance to the original ancestor for the generation."
)
```

## Last Generation

```{r, include = FALSE}
last_gen <-
x %>%
  filter(gen_num == max(gen_num))

dmat <- stringdistmatrix(last_gen$the_seq, method = 'hamming')
avg_hd <- mean(dmat)
dvec <- as.numeric(dmat)/min(nchar(last_gen$the_seq))
dd.f <- data.frame(dists = dvec)
```

The average pairwise HD is `r round(avg_hd, 2)` and normalized by the length of the sequence it is `r round(avg_hd/min(nchar(last_gen$the_seq)), 3)`.

```{r, results = 'asis'}
ggplot(dd.f, aes(x = dists)) + geom_density()
cat("\n\nFigure: Density plot of the pariwise distances in the last generation.\n\n")

deciles <-
data.frame(Decile = paste((0:10)*10, "%", sep = ''),
           Value = quantile(dvec, (0:10)/10))
row.names(deciles) <- NULL
cat("\n\nTable: The Deciles of the pairwise HDs.\n\n")
kable(deciles)
```
