---
title: First useful simulation.
output: 
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
      sharing:
        facebook: no
        twitter: no
      fontsettings:
        theme: sepia
---

```{r, include=FALSE}
library(yasss)
library(knitr)
library(rmarkdown)
library(ggthemr)

knitr::opts_chunk$set(echo = FALSE)
options(scipen = 99)
setCacheDir("/home/phillipl/projects/quasispecies_sim_reports/repo/quasispecies_sim_reports/first_useful_comparison/simpleCache")
```

```{r , include = FALSE}
source('../utilities/reporting_utilities.R')
```

# Simulation Description

Comparing populations with a fitness contraint to populations with no fitness constraint.

Simulate 10 datasets with 1024 individuals and uniformly assigned fitness scores.

Assign the first 1+2+4+8=15 sequences a fitness of 1 - to prevent massive parts of the dataset from getting excluded.

Then, do a fitness based selection with a score of 0.05.

Assign that as sim_id n to group 1.

Next select that same number of individuals from the last generation completely ignoring the fitness score and assign that as sim_id n+1 to group 2.

Repeat until 10 simulations in each dataset.

```{r, include=FALSE}

args1 <- list(
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 10,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 1/500)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL)
)

arg_sets <- list(`1` = list(group_label = "restricted",
                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and required fitness of 0.05",
                            args = args1),
                 `2` = list(group_label = "unrestricted",
                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and a random sample to match the size of the fitness restricted version",
                            args = args1))

group_membership <- NULL
all_sim_results <- NULL
i <- 0
for (i in 1:10){
#  x <- do.call(sim_pop, args1)
  x <- memoiseCache(fun = 'sim_pop', args = args1, cacheName = 'first_useful', seed = i)
  x$fitness_score[1:15] <- 1
  y <- get_fit_offspring(x, 0.1)
  original_col <- names(y)
  y$sim_id <- i*2-1
  y$group_label <- 'restricted'
  y <- y %>% select(sim_id, group_label, original_col)

  y_last_gen_size <- nrow(y %>% filter(gen_num == max(gen_num)))
  x_last_gen_size <- nrow(x %>% filter(gen_num == max(gen_num)))
  x1 <- (x %>% filter(gen_num == max(gen_num)))[sample(1:x_last_gen_size, y_last_gen_size),]
  x2 <- rbind((x %>% filter(gen_num != max(gen_num))), x1)
  x2_last_gen_size <- nrow(x2 %>% filter(gen_num == max(gen_num)))
  stopifnot(x2_last_gen_size == y_last_gen_size)

  x2$sim_id <- i*2
  x2$group_label <- "unrestricted"
  x2 <- x2 %>% select(sim_id, group_label, original_col)

  if (is.null(all_sim_results)){
    all_sim_results <- rbind(y, x2)
  } else {
    all_sim_results <- rbind(all_sim_results,
                             rbind(y, x2))
  }

  if (is.null(group_membership)){
    group_membership <- data.frame(sim_id = 1:2, group_id = 1:2)
  } else {
    group_membership <- rbind(group_membership,
      data.frame(sim_id = c(i*2-1, i*2), 
                 group_id = 1:2)
      )
  }

}
```

# Simulation Results

```{r test-child, child = '../utilities/comp_sim_chunk.Rmd'}
```
