---
title: First useful comparison
output: 
  html_document:
    toc: false
    theme: cerulean
---

```{r, include=FALSE}
library(yasss)
library(knitr)
library(rmarkdown)
library(tidyr)

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(echo = FALSE,
#                      base.dir = "/home/phillipl/projects/quasispecies_sim_reports/builds/first_useful_comparison/figures")
options(scipen = 99)
setCacheDir("/home/phillipl/projects/quasispecies_sim_reports/builds/first_useful_comparison/simpleCache")
report_seed <- 1000
set.seed(report_seed)

report_start_time <- proc.time()
```

```{r , include = FALSE}
source('../utilities/reporting_utilities.R')
```

```{r, eval = FALSE}
yasss:::restart_r()
```


# Description

Comparing populations with a fitness contraint to populations with no fitness constraint.

Simulate 10 datasets with 13 generations of size 2 producing 8192 individuals in the last generation and uniformly assigned fitness scores.

Assign perfect fitness to the first 5 generations to prevent massive parts of the dataset from getting excluded.

Then, do a fitness based selection with a score of 0.05.

Assign that as `sim_id` n to group 1.

Next select that same number of individuals from the last generation completely ignoring the fitness score and assign that as `sim_id` `n+1` to group 2.

Repeat until 10 simulations in each dataset.

```{r, include=FALSE}
chunk_seed <- 2000000

arg_set1 <- list(
  label = 'A',
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 10,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 1/500)),
  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
                           args = NULL),
  required_fitness = 0.1
)

arg_collection <- list(arg_set1)

many_pops <-
sim_proc_many_pops(arg_collection, n_sims = 10, fitness_processing =
                   'fit_unfit_pair')



#
#args1 <- list(
#  ancestors = paste(rep("A", 500), collapse = ''),
#  r0 = 2,
#  n_gen = 18,
#  n_pop = Inf,
#  mutator = list(fun = "mutator_uniform_fun",
#                 args = list(mu = 1/500)),
#  fitness_evaluator = list(fun = "fitness_evaluator_uniform_fun",
#                           args = NULL)
#)
#
#arg_sets <- list(`1` = list(group_label = "restricted",
#                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and required fitness of 0.1",
#                            args = args1),
#                 `2` = list(group_label = "unrestricted",
#                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and a random sample to match the size of the fitness restricted version",
#                            args = args1))
#
#group_membership <- data.frame(sim_id = numeric(0), 
#                               group_id = integer(0))
#
#all_sim_results <- data.frame(sim_id = numeric(0), 
#                              group_label = character(0), 
#                              gen_num = numeric(0), 
#                              id = numeric(0), 
#                              parent_id = numeric(0), 
#                              the_seq = character(0), 
#                              n_mut = numeric(0), 
#                              recomb_pos = numeric(0), 
#                              recomb_replaced = character(0), 
#                              recomb_partner = numeric(0), 
#                              recomb_muts = numeric(0), 
#                              fitness_score = numeric(0))
#
#sim_result <- sim_fit_unfit_pairs(args = arg_sets[[1]]$args, 
#                                  n_of_pairs = 10,
#                                  group1_label = arg_sets[[1]]$group_label,
#                                  group2_label = arg_sets[[2]]$group_label,
#                                  group1_id = 1, group2_id = 2,
#                                  all_sim_results = all_sim_results,
#                                  group_membership = group_membership,
#                                  seed_offset = report_seed + chunk_seed,
#                                  req_fitness = 0.1, 
#                                  n_perfect_fit = 31,
#                                  cacheName = 'first_useful',
#                                  verbose = TRUE)
#
#all_sim_results <- sim_result$all_sim_results
#group_membership <- sim_result$group_membership
#


#
#group_membership <- NULL
#all_sim_results <- NULL
#i <- 0
#for (i in 1:10){
##  x <- do.call(sim_pop, args1)
#  suppressMessages({
#  x <- memoiseCache(fun = 'sim_pop', args = args1, cacheName = 'first_useful', seed = i)
#  })
#  x$fitness_score[1:15] <- 1
#  y <- get_fit_offspring(x, 0.1)
#  original_col <- names(y)
#  y$sim_id <- i*2-1
#  y$group_label <- 'restricted'
#  y <- y %>% select(sim_id, group_label, original_col)
#
#  y_last_gen_size <- nrow(y %>% filter(gen_num == max(gen_num)))
#  x_last_gen_size <- nrow(x %>% filter(gen_num == max(gen_num)))
#  x1 <- (x %>% filter(gen_num == max(gen_num)))[sample(1:x_last_gen_size, y_last_gen_size),]
#  x2 <- rbind((x %>% filter(gen_num != max(gen_num))), x1)
#  x2_last_gen_size <- nrow(x2 %>% filter(gen_num == max(gen_num)))
#  stopifnot(x2_last_gen_size == y_last_gen_size)
#
#  x2$sim_id <- i*2
#  x2$group_label <- "unrestricted"
#  x2 <- x2 %>% select(sim_id, group_label, original_col)
#
#  if (is.null(all_sim_results)){
#    all_sim_results <- rbind(y, x2)
#  } else {
#    all_sim_results <- rbind(all_sim_results,
#                             rbind(y, x2))
#  }
#
#  if (is.null(group_membership)){
#    group_membership <- data.frame(sim_id = 1:2, group_id = 1:2)
#  } else {
#    group_membership <- rbind(group_membership,
#      data.frame(sim_id = c(i*2-1, i*2), 
#                 group_id = 1:2)
#      )
#  }
#
#}
```

# Simulation Results

```{r test-child, child = '../utilities/comp_sim_chunk.Rmd'}
```

# Total Running Time
```{r}
print(proc.time() - report_start_time)
```
