---
title: Basic homology based fitness
output: 
  bookdown::gitbook:
    config:
      toc:
        collapse: subsection
      sharing:
        facebook: no
        twitter: no
      fontsettings:
        theme: sepia
---

```{r, include=FALSE}
library(yasss)
library(knitr)
library(rmarkdown)
library(ggthemr)

knitr::opts_chunk$set(echo = FALSE,
                      base.dir = "/home/phillipl/projects/quasispecies_sim_reports/builds/basic_homolo_based_fitness/figures")
options(scipen = 99)
setCacheDir("/home/phillipl/projects/quasispecies_sim_reports/builds/basic_homolo_based_fitness/simpleCache")
set.seed(1)
source('../utilities/reporting_utilities.R')
```

```{r debugging-chunk, eval = FALSE}
yasss:::restart_r()
```

# Basic homology based fitness

## Description

A first basic report exploring the effects of assigning fitness based on homology.

Start with an ancestor of only As and then run two comparisons:

- One against a comparator that is only As and Xs - lets say 20% A - using the
  current fitness selection tools this will not work since the ancestor will
  die. Hack around it by assigning the first couple of generations perfect
  fitness. (currently the first 63 rows)
- One against a similat comparator, but with the As replaced with Cs

```{r, include=FALSE}
args1 <- list(
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 12,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 1/250)),
  fitness_evaluator = list(fun = "fitness_evaluator_homology_fun",
                           args = list(comparators = paste(rep('XXXXA', 100), collapse = ''),
                                       h2fs = "h2fs_univariate_linear_fun"))
)
args2 <- list(
  ancestors = paste(rep("A", 500), collapse = ''),
  r0 = 2,
  n_gen = 12,
  n_pop = Inf,
  mutator = list(fun = "mutator_uniform_fun",
                 args = list(mu = 1/250)),
  fitness_evaluator = list(fun = "fitness_evaluator_homology_fun",
                           args = list(comparators = paste(rep('XXXXC', 100), collapse = ''),
                                       h2fs = "h2fs_univariate_linear_fun"))
)


arg_sets <- list(`1` = list(group_label = "restricted_A",
                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and required fitness of 0.03 vs A",
                            args = args1),
                 `2` = list(group_label = "unrestricted_A",
                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and a random sample to match the size of the fitness restricted version vs A",
                            args = args1),
                 `3` = list(group_label = "restricted_C",
                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and required fitness of 0.03 vs C",
                            args = args2),
                 `4` = list(group_label = "unrestricted_C",
                            group_desc = "with r0 = 2, 10 generations and mu equal to 1/500 and a random sample to match the size of the fitness restricted version vs C",
                            args = args2))

ptm <- proc.time()
ltm <- proc.time()

group_membership <- NULL
all_sim_results <- NULL
i <- 1

library(profvis)
profvis({

for (i in 1:10){
#  x <- do.call(sim_pop, args1)
  print('===============================')
  print(i)
  print(proc.time() - ptm)
  print(proc.time() - ltm)
  ltm <- proc.time()
  suppressMessages({
  x <- memoiseCache(fun = 'sim_pop', args = args1, cacheName = 'basic_homolo', seed = i)
  })
  x$fitness_score[1:255] <- 1
  y <- get_fit_offspring(x, 0.02)
  original_col <- names(y)
  y$sim_id <- i*2-1
  y$group_label <- 'restricted_A'
  y <- y %>% select(sim_id, group_label, original_col)

  y_last_gen_size <- nrow(y %>% filter(gen_num == max(gen_num)))
  x_last_gen_size <- nrow(x %>% filter(gen_num == max(gen_num)))
  x1 <- (x %>% filter(gen_num == max(gen_num)))[sample(1:x_last_gen_size, y_last_gen_size),]
  x2 <- rbind((x %>% filter(gen_num != max(gen_num))), x1)
  x2_last_gen_size <- nrow(x2 %>% filter(gen_num == max(gen_num)))
  stopifnot(x2_last_gen_size == y_last_gen_size)

  x2$sim_id <- i*2
  x2$group_label <- "unrestricted_A"
  x2 <- x2 %>% select(sim_id, group_label, original_col)

  if (is.null(all_sim_results)){
    all_sim_results <- rbind(y, x2)
  } else {
    all_sim_results <- rbind(all_sim_results,
                             rbind(y, x2))
  }

  if (is.null(group_membership)){
    group_membership <- data.frame(sim_id = 1:2, group_id = 1:2)
  } else {
    group_membership <- rbind(group_membership,
      data.frame(sim_id = c(i*2-1, i*2), 
                 group_id = 1:2)
      )
  }

}

for (i in 11:20){
#  x <- do.call(sim_pop, args1)
  print('===============================')
  print(i)
  print(proc.time() - ptm)
  print(proc.time() - ltm)
  ltm <- proc.time()
  suppressMessages({
  x <- memoiseCache(fun = 'sim_pop', args = args2, cacheName = 'basic_homolo', seed = i)
  })
  x$fitness_score[1:255] <- 1
  y <- get_fit_offspring(x, 0.02)
  original_col <- names(y)
  y$sim_id <- i*2-1
  y$group_label <- 'restricted_C'
  y <- y %>% select(sim_id, group_label, original_col)

  y_last_gen_size <- nrow(y %>% filter(gen_num == max(gen_num)))
  x_last_gen_size <- nrow(x %>% filter(gen_num == max(gen_num)))
  x1 <- (x %>% filter(gen_num == max(gen_num)))[sample(1:x_last_gen_size, y_last_gen_size),]
  x2 <- rbind((x %>% filter(gen_num != max(gen_num))), x1)
  x2_last_gen_size <- nrow(x2 %>% filter(gen_num == max(gen_num)))
  stopifnot(x2_last_gen_size == y_last_gen_size)

  x2$sim_id <- i*2
  x2$group_label <- "unrestricted_C"
  x2 <- x2 %>% select(sim_id, group_label, original_col)

  if (is.null(all_sim_results)){
    all_sim_results <- rbind(y, x2)
  } else {
    all_sim_results <- rbind(all_sim_results,
                             rbind(y, x2))
  }

  if (is.null(group_membership)){
    group_membership <- data.frame(sim_id = 1:2, group_id = 1:2)
  } else {
    group_membership <- rbind(group_membership,
      data.frame(sim_id = c(i*2-1, i*2), 
                 group_id = 3:4)
      )
  }
}

})

```

## Simulation Results

```{r test-child, child = '../utilities/comp_sim_chunk.Rmd'}
```
